---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/", # nolint
  eval = TRUE
)
```

# epinow2cast: EpiNow2 powered by epinowcast

> **This is an experimental rewrite of [EpiNow2](https://github.com/epiforecasts/EpiNow2) that replaces all internal Stan models with calls to the [epinowcast](https://package.epinowcast.org/) package.** It is not yet released and is under active development. See the [feature status](#feature-status) below.

## What is this?

`{epinow2cast}` provides the familiar EpiNow2 interface (`estimate_infections()`, `epinow()`, `regional_epinow()`) but delegates all Bayesian inference to `{epinowcast}`. This eliminates ~9,000 lines of Stan code from EpiNow2 and unifies the two packages under a single computational backend.

## Feature status

| Feature | Status | Notes |
|---------|--------|-------|
| `estimate_infections()` | Working | Renewal equation via epinowcast |
| `estimate_truncation()` | Working | Requires [epinowcast PR #756](https://github.com/epinowcast/epinowcast/pull/756) |
| `simulate_infections()` | Working | Pure R implementation |
| `epinow()` | Working | |
| `regional_epinow()` | Working | |
| `get_samples()` | Working | R, infections, growth_rate, reported_cases |
| `get_predictions()` | Working | summary, sample, quantile formats |
| `get_parameters()` | Working | Returns generation time and delays |
| `summary()` / `plot()` | Working | |
| Random walk on Rt | Working | `rt_opts(rw = 7)` for weekly |
| Day-of-week effects | Working | `obs_opts(week_effect = TRUE)` |
| Forecasting | Working | `forecast_opts(horizon = N)` |
| Poisson / NegBin | Working | `obs_opts(family = ...)` |
| Reporting delays | Working | Via `latent_reporting_delay` |
| Gaussian process | Removed | Use `rt_opts(rw = ...)` instead |
| Back-calculation | Removed | |
| Population depletion | Removed | |
| `estimate_secondary()` | Removed | Use epinowcast directly |
| `forecast_infections()` | Removed | Use `forecast_opts()` in `estimate_infections()` |
| rstan backend | Removed | cmdstanr only (via epinowcast) |

## Installation

This package is not on CRAN. Install from this repository:

```{r, eval = FALSE}
# install.packages("pak")
pak::pkg_install("sbfnk/epinow2cast@epinowcast-backend")
```

Requires [`cmdstanr`](https://mc-stan.org/cmdstanr/) and a working CmdStan installation. See the [cmdstanr documentation](https://mc-stan.org/cmdstanr/articles/cmdstanr.html) for setup instructions.

## Resources

<details> <summary> Getting Started </summary>

The Getting Started vignette (see `vignette("EpiNow2")`)
is your quickest entry point to the package. It provides a quick run through of
the two main functions in the package and how to set up them up. It also
discusses how to summarise and visualise the results after running the models.

More broadly, users can also learn the details of estimating delay distributions, nowcasting, and forecasting in a structured way through the free and open short-course, ["Nowcasting and forecasting infectious disease dynamics"](https://nfidd.github.io/nfidd/), developed by some authors of this package.

</details>

<details> <summary> Package website </summary>

The package has two websites: one for
[the stable release version on CRAN](https://epiforecasts.io/EpiNow2/), and
another for [the version in development](https://epiforecasts.io/EpiNow2/dev/).
These two provide various resources for learning about the package, including
the function reference, details about each model (model definition), workflows
for each model (usage), and case studies or literature of applications of
the package. However, the development website may contain experimental features
and information not yet available in the stable release.

</details>

<details> <summary> End-to-end workflows </summary>

The workflow vignette (see `vignette("estimate_infections_workflow")`)
provides guidance on the end-to-end process of estimating reproduction
numbers and performing short-term forecasts for a disease spreading in a

</details>

<details> <summary> Model definitions </summary>

In different vignettes we provide the mathematical definition of each model.
For example, the model definition vignette for `estimate_infections()` can be
found in `vignette("estimate_infections")`.

</details>

<details> <summary> Example implementations </summary>

A simple example of using the package to estimate a national Rt for Covid-19 can be found [here](https://gist.github.com/seabbs/163d0f195892cde685c70473e1f5e867).

</details>

## Contributing

We welcome all contributions. If you have identified an issue with the package,
you can file an issue [here](https://github.com/epiforecasts/EpiNow2/issues). We also welcome additions and extensions to the underlying model either in the form of options or improvements. If you wish to contribute in any form, please follow the
[package contributing guide](https://github.com/epiforecasts/EpiNow2/blob/main/.github/CONTRIBUTING.md).

## Contributors
















<!-- ALL-CONTRIBUTORS-LIST:START - Do not remove or modify this section -->
<!-- prettier-ignore-start -->
<!-- markdownlint-disable -->

All contributions to this project are gratefully acknowledged using the [`allcontributors` package](https://github.com/ropensci/allcontributors) following the [allcontributors](https://allcontributors.org) specification. Contributions of any kind are welcome!

### Code


<a href="https://github.com/epiforecasts/EpiNow2/commits?author=seabbs">seabbs</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=sbfnk">sbfnk</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=jamesmbaazam">jamesmbaazam</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=joeHickson">joeHickson</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=hsbadr">hsbadr</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=pitmonticone">pitmonticone</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=actions-user">actions-user</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=ellisp">ellisp</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=kaitejohnson">kaitejohnson</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=jdmunday">jdmunday</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=pearsonca">pearsonca</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=Bisaloo">Bisaloo</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=JAllen42">JAllen42</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=adamkucharski">adamkucharski</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=adrian-lison">adrian-lison</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=avehtari">avehtari</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=andrjohns">andrjohns</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=claude">claude</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=jcken95">jcken95</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=LloydChapman">LloydChapman</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=medewitt">medewitt</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=nikosbosse">nikosbosse</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=sophiemeakin">sophiemeakin</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/commits?author=zsusswein">zsusswein</a>



### Issue Authors


<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Araulfernandezn">raulfernandezn</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Apcarbo">pcarbo</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Ajohnaponte">johnaponte</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Asophie-schiller">sophie-schiller</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Amunozedg">munozedg</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Akathsherratt">kathsherratt</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Ayungwai">yungwai</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Akgostic">kgostic</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Afkrauer">fkrauer</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Aphilturk">philturk</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Akrageth">krageth</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Atony352">tony352</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Ausername-rp">username-rp</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3AHAKGH">HAKGH</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3AAndrewRiceMGW">AndrewRiceMGW</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Abrynhayder">brynhayder</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3ARichardMN">RichardMN</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Aandrybicio">andrybicio</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Arhamoonga">rhamoonga</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Afurqan915">furqan915</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3AMFZaini1984">MFZaini1984</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Afabsig">fabsig</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Aaffans">affans</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3AGauriSaran">GauriSaran</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Adavidvilanova">davidvilanova</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Ajrcpulliam">jrcpulliam</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Adajmcdon">dajmcdon</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Ajoshwlambert">joshwlambert</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Aavallecam">avallecam</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Aathowes">athowes</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Alorenzwalthert">lorenzwalthert</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Anlinton">nlinton</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Amartinamcm">martinamcm</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Ajonathonmellor">jonathonmellor</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3ATimTaylor">TimTaylor</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Aciaramccarthy1">ciaramccarthy1</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3ASamuelBrand1">SamuelBrand1</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Adamonbayer">damonbayer</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Avalentinedwv">valentinedwv</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Acoderabbitai">coderabbitai</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3AHenrikBengtsson">HenrikBengtsson</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+author%3Abob-carpenter">bob-carpenter</a>



### Issue Contributors


<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3Ajhellewell14">jhellewell14</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3Athlytras">thlytras</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3ALizaHadley">LizaHadley</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3Antorresd">ntorresd</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3Amicahwiesner67">micahwiesner67</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3Apaigemiller">paigemiller</a>, 
<a href="https://github.com/epiforecasts/EpiNow2/issues?q=is%3Aissue+commenter%3AWardBrian">WardBrian</a>


<!-- markdownlint-enable -->
<!-- prettier-ignore-end -->
<!-- ALL-CONTRIBUTORS-LIST:END -->
















